---
title: "REV vs REV_TAX vs DMSO"
author: "Fabio Bedin | MS-Unit"
date: "12 luglio 2021"
output: html_document
---

```{r librerie, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.align = "center", warning=FALSE, message=FALSE, fig.height=10, fig.width=10)
library("MBQN")
library("dplyr")
library("tidyr")
library("DEP")
library("SummarizedExperiment")
library("preprocessCore")
library("tibble")
library("ggplot2")
library("enrichR")
library("DT")
library("stringr")
library("patchwork")
library("here")
library("datapasta")
library("ggstatsplot")
library("UpSetR")
library("ggrepel")
library("visdat")
library("naniar")
library("gprofiler2")
library("openxlsx")
```

```{r Custom Functions}

source(here::here("code/custom_functions.R"))

```

```{r}
data <- read.csv(here::here("data/proteinGroups_DMSO_REV_REV_TAX.txt"), header = TRUE,stringsAsFactors = FALSE, sep = "\t")

data <- data[data$Reverse != "+" & data$Potential.contaminant != "+" & data$Only.identified.by.site != "+" & data$Peptides >= 2,]

data_unique <- make_unique(data, "Gene.names", "Protein.IDs", delim = ";")
```

# PCA all samples

```{r}
expdesign <- read.table(here::here("data/expdesign.tsv"), header = T, stringsAsFactors = F)

columns<-match(paste("LFQ.intensity.",expdesign$label,sep=""),colnames(data_unique))

data_se <- make_se(data_unique, columns, expdesign)

data_filt <- filter_missval(data_se, thr = 0)

data_diff <- test_diff_BH(data_filt, type = "all")

dep <- add_rejections(data_diff, alpha = 0.05, lfc = 1)

results<-get_results(dep)

plot_pca(dep, x = 1, y = 2, n = 500, point_size = 4, indicate = c("condition", "replicate"))

```


```{r define_cond_1}
cond_1 <- "REV."

cond_2 <- "DMSO"

test<- paste(cond_1,cond_2,sep="_vs_")
```

# **Analysis** {.tabset .tabset-fade}

## **`r test`** {.tabset .tabset-fade}

```{r expdes_1}
expdesign <- read.table(here::here("data/expdesign.tsv"), header = T, stringsAsFactors = F)

conditions<-c(cond_1,cond_2)

expdesign <- subset(expdesign, condition %in% conditions)

columns<-match(paste("LFQ.intensity.",expdesign$label,sep=""),colnames(data_unique))

data_se <- make_se(data_unique, columns, expdesign)

data_filt_Unique <- filter_missval(data_se, thr = 0)

data_filt <- filter_by_grouping(data_se, thr = 0)
```

### Proteins

```{r proteins_1}
plot_numbers_lables(data_filt_Unique)
```

### normalization

```{r normalization_1}
plot_normalization(data_filt_Unique)
```

### frequency

```{r frequency_1}
plot_frequency(data_filt_Unique)
```

### upset plot {.tabset .tabset-fade}

#### `r cond_1`

```{r upset_1.1}
assay(data_filt_Unique) %>% 
  as.data.frame() %>% 
  select(starts_with(cond_1)) %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 3,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = "coral2",
        main.bar.color  = "gray44")
```

#### `r cond_2`

```{r upset_1.2}
assay(data_filt_Unique) %>% 
  as.data.frame() %>% 
  select(starts_with(cond_2)) %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 3,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = "turquoise3",
        main.bar.color  = "gray44")
```

#### Both

```{r upset_1.3}
assay(data_filt_Unique) %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 6,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = rep(c("turquoise3", "coral2"), each = 3),
        main.bar.color  = "gray44")
```


### DEP analysis {.tabset .tabset-fade}

```{r diff_1}
data_diff <- test_diff_BH(data_filt, type = "manual", test = test)

dep <- add_rejections(data_diff, alpha = 0.05, lfc = 1)

results<-get_results(dep)
```

Significant: `r results %>% filter(significant) %>% nrow()`

#### PCA

```{r PCA_1}
plot_pca(dep, x = 1, y = 2, n = 500, point_size = 4, indicate = c("condition", "replicate"))
```

#### Volcano 

```{r volcano_1}
plot_volcano(dep, contrast=test, add_names=T,label_size=5,adjusted = F)
```

### Unique

```{r unique_1}
unique_pair(data_filt_Unique, conditions = conditions)
```

### CC Unique

```{r}
tab_unique <- unique_pair(data_filt_Unique, conditions = conditions, table = T)

GO_name <- tab_unique %>% filter(unique_REV.) %>% pull(name)

GO <- gost(query = GO_name, organism = "hsapiens", sources = "GO:CC", ordered_query = F, significant = F)

gostplot(GO, capped = TRUE, interactive = TRUE)
```


```{r define_cond_2}
cond_1 <- "REV."

cond_2 <- "REVTAX"

test<- paste(cond_1,cond_2,sep="_vs_")
```


## **`r test`** {.tabset .tabset-fade}

```{r expdes_2}
expdesign <- read.table(here::here("data/expdesign.tsv"), header = T, stringsAsFactors = F)

conditions<-c(cond_1,cond_2)

expdesign <- subset(expdesign, condition %in% conditions)

columns<-match(paste("LFQ.intensity.",expdesign$label,sep=""),colnames(data_unique))

data_se <- make_se(data_unique, columns, expdesign)

data_filt_Unique <- filter_missval(data_se, thr = 0)

data_filt <- filter_by_grouping(data_se, thr = 0)
```

### Proteins

```{r proteins_2}
plot_numbers_lables(data_filt_Unique)
```

### normalization

```{r normalization_2}
plot_normalization(data_filt_Unique)
```

### frequency

```{r frequency_2}
plot_frequency(data_filt_Unique)
```

### upset plot {.tabset .tabset-fade}

#### `r cond_1`

```{r upset_2.1}
assay(data_filt_Unique) %>% 
  as.data.frame() %>% 
  select(starts_with(cond_1)) %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 3,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = "coral2",
        main.bar.color  = "gray44")
```

#### `r cond_2`

```{r upset_2.2}
assay(data_filt_Unique) %>% 
  as.data.frame() %>% 
  select(starts_with(cond_2)) %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 3,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = "turquoise3",
        main.bar.color  = "gray44")
```

#### Both

```{r upset_2.3}
assay(data_filt_Unique) %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 6,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = rep(c("turquoise3", "coral2"), each = 3),
        main.bar.color  = "gray44")
```


### DEP analysis {.tabset .tabset-fade}

```{r diff_2}
data_diff <- test_diff_BH(data_filt, type = "manual", test = test)

dep <- add_rejections(data_diff, alpha = 0.05, lfc = 1)

results<-get_results(dep)
```

Significant: `r results %>% filter(significant) %>% nrow()`

#### PCA

```{r PCA_2}
plot_pca(dep, x = 1, y = 2, n = 500, point_size = 4, indicate = c("condition", "replicate"))
```

#### Volcano 

```{r volcano_2}
plot_volcano(dep, contrast=test, add_names=T,label_size=5,adjusted = F)
```

### Unique

```{r unique_2}
unique_pair(data_filt_Unique, conditions = conditions)
```

### CC Unique

```{r}
tab_unique <- unique_pair(data_filt_Unique, conditions = conditions, table = T)

GO_name <- tab_unique %>% pull(name)

GO <- gost(query = GO_name, organism = "hsapiens", sources = "GO:CC", ordered_query = F, significant = F)

gostplot(GO, capped = TRUE, interactive = TRUE)
```

```{r define_cond_3}
cond_1 <- "REVTAX"

cond_2 <- "DMSO"

test<- paste(cond_1,cond_2,sep="_vs_")
```


## **`r test`** {.tabset .tabset-fade}

```{r expdes_3}
expdesign <- read.table(here::here("data/expdesign.tsv"), header = T, stringsAsFactors = F)

conditions<-c(cond_1,cond_2)

expdesign <- subset(expdesign, condition %in% conditions)

columns<-match(paste("LFQ.intensity.",expdesign$label,sep=""),colnames(data_unique))

data_se <- make_se(data_unique, columns, expdesign)

data_filt_Unique <- filter_missval(data_se, thr = 0)

data_filt <- filter_by_grouping(data_se, thr = 0)
```

### Proteins

```{r proteins_3}
plot_numbers_lables(data_filt_Unique)
```

### normalization

```{r normalization_3}
plot_normalization(data_filt_Unique)
```

### frequency

```{r frequency_3}
plot_frequency(data_filt_Unique)
```

### upset plot {.tabset .tabset-fade}

#### `r cond_1`

```{r upset_3.1}
assay(data_filt_Unique) %>% 
  as.data.frame() %>% 
  select(starts_with(cond_1)) %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 3,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = "coral2",
        main.bar.color  = "gray44")
```

#### `r cond_2`

```{r upset_3.2}
assay(data_filt_Unique) %>% 
  as.data.frame() %>% 
  select(starts_with(cond_2)) %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 3,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = "turquoise3",
        main.bar.color  = "gray44")
```

#### Both

```{r upset_3.3}
assay(data_filt_Unique) %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 6,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = rep(c("turquoise3", "coral2"), each = 3),
        main.bar.color  = "gray44")
```


### DEP analysis {.tabset .tabset-fade}

```{r diff_3}
data_diff <- test_diff_BH(data_filt, type = "manual", test = test)

dep <- add_rejections(data_diff, alpha = 0.05, lfc = 1)

results<-get_results(dep)
```

Significant: `r results %>% filter(significant) %>% nrow()`

#### PCA

```{r PCA_3}
plot_pca(dep, x = 1, y = 2, n = 500, point_size = 4, indicate = c("condition", "replicate"))
```

#### Volcano 

```{r volcano_3}
plot_volcano(dep, contrast=test, add_names=T,label_size=5,adjusted = F)
```

### Unique

```{r unique_3}
unique_pair(data_filt_Unique, conditions = conditions)
```

### CC Unique

```{r}
tab_unique <- unique_pair(data_filt_Unique, conditions = conditions, table = T)

GO_name <- tab_unique %>% filter(unique_REVTAX) %>% pull(name)

GO <- gost(query = GO_name, organism = "hsapiens", sources = "GO:CC", ordered_query = F, significant = F)

gostplot(GO, capped = TRUE, interactive = TRUE)
```

# **Unique in each conditions**

```{r super_unique}
expdesign <- read.table(here::here("data/expdesign.tsv"), header = T, stringsAsFactors = F)

columns<-match(paste("LFQ.intensity.",expdesign$label,sep=""),colnames(data_unique))

data_se <- make_se(data_unique, columns, expdesign)

data_filt <- filter_missval(data_se, thr = 0)

assay(data_filt) %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>%
  mutate(intensity = case_when(is.na(intensity) ~ 0, TRUE ~ as.numeric(1))) %>%
  group_by(rowname) %>%
  mutate(unique = if_else(sum(intensity) == 3, T, F)) %>%
  ungroup() %>%
  filter(unique) %>% 
  mutate(condition = gsub(pattern = "..$", replacement = "",x = samples)) %>% 
  group_by(rowname, condition) %>% 
  mutate(unique_2 = if_else(sum(intensity) == 0, F, T)) %>% 
  ungroup() %>% 
  mutate(unique_3 = case_when(unique_2 ~ condition)) %>% 
  select(rowname, samples, intensity, unique_3) %>%
  group_by(rowname) %>% 
  fill(unique_3, .direction = "downup") %>% 
  ungroup() %>% 
  pivot_wider(id_cols = c(rowname, unique_3) ,names_from = samples, values_from = intensity) %>% 
  dplyr::rename(Unique = unique_3, name = rowname) %>% 
  select(1:2) %>%
  as.data.frame() %>% 
    DT::datatable(options = list(
      columnDefs = list(list(className = 'dt-center', targets = 1:2)),
      pageLength = 10))
```


## {-}



